Research Notes

### 2018-10-15 writeups/notes.tex
check in 2d_1_gradual param scan, mostly working

This is because we performed the naive substitution $k, \omega$ from the parent
wave. Instead, it is advisable to scale $k, \omega$ to the same $k_{turb}$
length scale. We make ansatz that the cascade is roughly Kolmogorov, and so
(TODO work this out) we obtain that we should choose $\frac{\omega_0}{k_z^2 \nu}
\sim \p*{\frac{k_{turb}}{k_z}}^{4/3}$; this comes from scaling $\omega, k$ per
the Kolmogorov prescription $\omega_l u_l^2$ is approximately constant over
length scale $l$.

@@

\section{Validity of the Linear Approximation}

The real physics is expected to deviate from our computation near the critical
layer owing primarily to two effects: nonlinear advective terms and diffusivity.
Hazel 1967 found numerically that including diffusive tersm in the equations of
motion does not change the attenuation factor; even in the presence of
dissipation no reflection/transimission are observed. This makes sense;
examining the equations of motion, the effect of an extra $\pd{}{t} \to \pd{}{t}
- \nu \nabla^2$ is to take $\omega_{cr} \to \omega_{cr} + i\nu k_x^2 - i\nu
\ptd{}{z}$, giving the singularity at $\omega_{cr} + i\nu k_x^2 = 0$ the
positive imaginary component we before asserted.

On the other hand, other studies e.g.\ Brown \& Stewartson 1980, Winters \&
d'Asaro 1994, find that once the wave amplitude becomes nonnegligible, sizable
reflection can be found. In contrast, Ogilvie \& Barker 2010 find that in
geometrically focused breaking, no reflection is observed, despite the nonlinear
effects being stronger than the dissipitive ones. The study of how nonlinear and
diffusive effects compete is then nontrivial.

>>>>

This is because we performed the naive substitution $k, \omega$ from the parent
wave. Instead, it is advisable to scale $k, \omega$ to the same $k_{turb}$
length scale. We first make ansatz that the cascade is roughly Kolmogorov, in
which case we do the hand-wavy argument that energy $u_l^2$ at a certain length
scale $l$ is dissipated on timescale $\omega_l = u_l / l$, and so the total
energy dissipated from the original flow $\epsilon = u_l^3/l$. Thus, $\nu k_l^2
\sim u_lk_l$ gives $\frac{\omega_0}{\nu k_0^2}\p*{\frac{k_0}{k_{turb}}}^{4/3} =
1$. This isn't quite right though, as per comparison.

My hypothesis is currently two fold: first, the turbulent cascade is not
instantaneous, so we can solve a diffusion equation in $k$ space to figure out
how far the wave can propagate (and grow in amplitude) before the Kolmogorov
spectrum is obtained. Second, the turbulence causes some energy to be reflected
in higher $k$ modes, which interact nonlinearly with the flow at lower
amplitudes.

@@

\section{Competition between Viscous and Nonlinear Forces}

As we zoom in on the critical layer, we expect to be able to take Boussinesq
approximation $H \to \infty$ within the immediate vicinity of the critical
layer.

In this limit, the group velocity is just $v_{g,z} = -\frac{Nk_xk_z}{\p*{k_x^2 +
k_z^2}^{3/2}} = -\omega\frac{k_z}{k_x^2 + k_z^2}$. As we approach corotation,
$\omega \to 0, \omega \approx \frac{Nk_x}{k_z}, k_z \to \infty$, and so $v_{g,
z} \to 0$. Then considering an energy flux incident on the critical layer
clearly shows that energy must accumulate near the critical layer, and must be
dissipated somehow. This must be either via viscous forces or nonlinear
cascades.

The characteristic strength of the viscous force $\nu \nabla^2$ is $\nu k_z^2$,
while the nonlinear advective term has characteristic strength $u_{1z}k_z$. Note
that in the Boussinesq approximation $\abs*{u_{1z}}$ is constant during
propagation. Thus, the ratio of forces $R = \frac{f_{\nu}}{f_{adv}} \sim
\frac{\nu k_z}{u_{1z}}$. Note that this ratio varies spatially. Qualitatively,
$R$ is important since it dictates whether the wave is damped viscously before
it needs to be treated in nonlinear fashion or whether a full
nonlinear/hydrodynamic treatment is necessary.

We want to understand for what viscosity we expect linear behavior to dominate.
Thus, at critical $z_{c}$ where nonlinear effects become important, we want $R
\gtrsim 1$. Shifting to coordinate system $\delta z$ where $\delta z = 0$ is the
corotation resonance, we can write $\abs*{\omega_{cr}} = k_xu_0' \delta z$ where
$u_0' = \at{\pd{u_0}{z}}_{\delta z = 0}$, and $k_z = \frac{N}{u_0' \delta z}$.
Then the breaking criterion is $\frac{u_{1z}k_z(z)}{\omega_{cr}(z)} \sim 1$, or
$\delta z_{c} \sim \sqrt{\frac{u_{1z}N}{k_xu_0'^2}}$. We can then evaluate
\begin{equation}
    R\p*{z_{c}} = \frac{\nu k_z}{u_{1z}} = \nu \sqrt{\frac{Nk_x}{u_{1z}^3}}.
\end{equation}

It is helpful also to write this down for any arbitrary hyperviscosity $f_{visc}
= \nu^{(n)} \nabla^{(n)}$, which is cited without proof
\begin{equation}
    R\p*{z_c} = \nu^{(n)} \sqrt{\frac{N^{n - 1}k_x^{n - 1}}{u_{1z}^{n + 1}}}.
\end{equation}
